---
title: "`kiwidecoder`: A Unified Pipeline for Barcode and QR-Code Decoding in Clonal Trials"
author: "Kathy Ruggiero"
date: "2026-01-29"
output: rmarkdown::html_vignette
appendix: yes
vignette: >
  %\VignetteIndexEntry{End-to-End Workflow and Diagnostics in kiwidecoder}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1. Introduction

Clonal trial imaging at scale produces **thousands of photos** spread across a complex directory tree on a network drive. Each image typically contains one or more **barcodes or QR codes** that encode:

*   plant genotypes,
*   scion identifiers,
*   locations,
*   and other trial metadata.

The goal of `kiwidecoder` is to provide a **unified, traceable, and robust pipeline** that:

1.  Scans a root directory on a (slow, network-mounted) file system,
2.  Builds **per-directory Excel indexes** of image files,
3.  Decodes **all barcodes and QR codes** in those images using ZXing-CPP (via Python),
4.  Resolves decoded values across **sequentially numbered filenames**, and
5.  Writes a **joined long-format table** back into each Excel file for downstream analysis.

In addition, the package provides:

*   low-level functions for **single-image decoding**,
*   diagnostic tools for **difficult images**,
*   helpers for **Python/ZXing troubleshooting**, and
*   mechanisms for running long scans in the **background**.

This vignette explains:

*   how barcode detection works under the hood,

*   how to set up the Python environment,

*   the end-to-end workflow:

    ```text
    scan_directories()
          ↓
     write_directory_index()  (per-directory Excel index)
          ↓
     resolve_folder_sequence()
          ↓
     joined index + barcode results per Excel
    ```

*   and how to use diagnostic tools when images fail to decode.

# 2. How Barcode Detection Works in `kiwidecoder`

*(A didactic explanation for users who may be new to R, Python, or barcode decoding)*

The functions in `kiwidecoder` look like ordinary R code, but the actual barcode decoding relies on **Python** and **ZXing-CPP**, a high-performance C++ barcode engine.

There are two main conceptual layers:

1.  **Single-image decoding**  
    `detect_codes_all_zxing(path)`  
    → opens one image via Python’s `PIL.Image`, runs ZXing-CPP, and returns one row per detected symbol.

2.  **Folder / index-based workflows**
    *   Historically: `scan_one_folder()` for simple, non-recursive tests.
    *   Currently:  
        `scan_directories()` + `write_directory_index()` + `resolve_folder_sequence()`  
        for high-throughput decoding and Excel integration.

The single-image decoder is the **core engine**; everything else orchestrates it across many files and manages the resulting tables.

## 2.1 Python, Pillow, and ZXing-CPP

Barcode detection is performed using the Python bindings to **ZXing-CPP**, a C++ library capable of decoding many 1D and 2D barcode formats, including:

*   PDF417
*   QR Code
*   Code39 / Code128
*   DataMatrix (with constraints in this workflow)

R does not currently have an equivalent native library, so `kiwidecoder` uses:

*   **`reticulate`** – to call Python from R,
*   **`Pillow` (PIL)** – to load images into Python,
*   **`zxingcpp`** – to decode barcodes from the Python image object.

Users do **not** need to write any Python code; the R functions manage all calls into Python. The only requirement is that R has access to a Python environment with `pillow` and `zxingcpp` installed.

This applies both to:

*   local R / RStudio sessions, and
*   Posit Workbench sessions on shared servers.

## 2.2 Conceptual data flow per image

For a single image, the data flow looks like:

```text
JPEG/PNG file on disk
        ↓
normalizePath() in R
        ↓
PIL.Image.open() in Python
        ↓
zxingcpp.read_barcodes()  (ZXing-CPP engine)
        ↓
list of Python result objects
        ↓
extract format + text → R vectors
        ↓
tibble: one row per symbol, plus classification
```

The function guarantees **at least one row per image** by returning a canonical NA row when:

*   the image cannot be opened,
*   Python modules are missing,
*   ZXing finds no symbol candidates,
*   or only unsupported formats (e.g., DataBar) are encountered.

This property is crucial for downstream **traceability** (see Section 4.4).

# 3. Setting Up the Barcode-Decoding Environment

## 3.1 Why Python is needed

R does not include a high-quality native decoder for formats like PDF417 or Code 128. `kiwidecoder` therefore uses `reticulate` to access Python and load:

*   **`Pillow`** (a.k.a. `PIL`) – opens and manipulates images,
*   **`zxingcpp`** – the barcode-decoding engine.

You do **not** write Python; you simply ensure these packages are installed once.

***

## 3.2 Step 1 — Install `reticulate` (once)

```r
install.packages("reticulate")
```

***

## 3.3 Step 2 — Install required Python libraries (once)

```r
reticulate::py_install(
  packages = c("pillow", "zxingcpp"),
  pip = TRUE
)
```

This will:

*   create a Python environment if needed,
*   install `pillow`,
*   install `zxingcpp`,
*   register the environment for future R sessions.

***

## 3.4 Step 3 — Verify that R can see Python

```r
reticulate::py_config()
```

Check that:

*   a Python executable is detected, and
*   `pillow` and `zxingcpp` are available in that environment.

Once this setup is completed, you can run decoding functions directly:

```r
detect_codes_all_zxing("path/to/image.jpg")
```

## 3.5 Notes for Posit Workbench users

On Posit Workbench:

*   In many setups, `py_install()` can create a **user-specific** environment in your home directory.
*   If permissions are restricted, your admin may need to install:

    *   `pillow`
    *   `zxingcpp`

Once they exist in the environment used by your R sessions, `kiwidecoder` will work normally.

# 4. The End-to-End High-Throughput Workflow

This section describes the **official production pipeline** for large image sets on a network-mounted drive.

## 4.1 High-level workflow overview

We assume you have a root directory like:

```text
N:/Projects/Stage2 Clonal Trials/Images/
    ├── Male Photos 2024_25/
    │     ├── IMG_0001.JPG
    │     ├── IMG_0002.JPG
    │     └── ...
    ├── Greenhouse/
    │     ├── bench_001.JPG
    │     └── ...
    └── ...
```

The workflow is:

```text
(1) scan_directories()
      ↓
Per-directory Excel indexes via write_directory_index():
   N:/.../Male Photos 2024_25/Male Photos 2024_25.xlsx
   N:/.../Greenhouse/Greenhouse.xlsx
   ...

(2) resolve_folder_sequence()
      ↓
Joined index + barcode tables written back to those Excel files:
   - original columns (dir, subdir, file_name, ext, rel_path, ...)
   - barcode columns (index, code_format, text, type, prefix, number, resolved_text)
```

**Key design choices:**

*   Network I/O is slow → we use **depth-first traversal + checkpointing**.
*   Excel files must not be corrupted → we use **safe atomic writes**.
*   Every image must be traceable → we guarantee **at least one row per image** at all stages.
*   Filenames encode useful structure → we use `resolve_sequence()` to propagate meaning.

## 4.2 Step 1 — Scanning and indexing with `scan_directories()`

`scan_directories()`:

*   walks the directory tree under `root`,
*   filters files by extension and optional filename prefixes,
*   calls `write_directory_index()` for each directory,
*   writes an Excel index per directory,
*   tracks progress via a checkpoint, so long runs can be resumed.

```{r scan-directories, eval = FALSE}
library(kiwidecoder)   # replace with actual package name

# Set the root of your image tree on the network drive.
image_root <- "N:/path/to/your/image/root"  # <-- EDIT THIS

scan_result <- scan_directories(
  root            = image_root,
  extensions      = c("jpg", "jpeg", "png"),
  prefixes        = c("^PMC_[0-9]{5}$", "^DSC_[0-9]{4}$",
                      "^IMG_[0-9]{4}$", "^P[0-9]{7}$"),
  checkpoint_file = "scan_checkpoint.rds",
  resume          = TRUE,
  dry_run         = FALSE,
  log_file        = file.path(image_root, "scan_log.csv"),
  max_path_length = 255
)

str(scan_result$dirs_scanned)
head(scan_result$long_path_warnings)
```

### When to use `dry_run = TRUE`

If you are unsure:

*   how many directories will be visited,
*   whether path lengths are safe for Excel,
*   whether your `prefixes` are correct,

run a dry run first:

```{r scan-dry-run, eval = FALSE}
scan_directories(
  root            = image_root,
  dry_run         = TRUE,
  log_file        = file.path(image_root, "scan_log.csv")
)
```

This scans, logs, and updates the checkpoint, but does **not** write Excel files.

This is particularly useful on **network drives** where scanning can take many hours.

## 4.3 Step 2 — Per-directory indexing with `write_directory_index()`

For each directory discovered, `scan_directories()` calls:

```r
write_directory_index(
  dir        = current_dir,
  extensions = extensions,
  prefixes   = prefixes,
  excel_path = excel_path
)
```

This produces a simple per-directory Excel file with one row per image and columns like:

*   `dir`      – full directory path containing the image,
*   `subdir`   – basename of `dir`,
*   `file_name`– the image filename,
*   `ext`      – lowercased extension (`"jpg"`, `"png"`, ...),
*   `rel_path` – relative path within `dir` (by default the same as `file_name`).

You can also use it directly for a single folder:

```{r write-directory-index-single, eval = FALSE}
one_dir <- file.path(image_root, "ExampleSubdir")  # <-- EDIT
index_path <- file.path(one_dir, "ExampleSubdir.xlsx")

index_df <- write_directory_index(
  dir        = one_dir,
  extensions = c("jpg", "jpeg", "png"),
  prefixes   = NULL,  # index all matching images
  excel_path = index_path
)

index_df
```

## 4.4 Data lineage and traceability

A key design principle in `kiwidecoder` is **per-image traceability**:

*   Every image in the directory tree appears in exactly one per-directory Excel index (subject to extension/prefix filters).
*   Every decoding step produces **at least one row per image**:
    *   `detect_codes_all_zxing()` returns an NA row if needed.
    *   `resolve_folder_sequence()` joins decoding results back to original index rows.
*   The final joined Excel tables contain:
    *   original structural columns (`dir`, `subdir`, `file_name`, `ext`, `rel_path`, `full_path`, etc.), plus
    *   per-symbol decoding data (`index`, `code_format`, `text`, `type`, `resolved_text`, etc.).

This means you can always:

*   trace any decoded value back to:
    *   its **source image file**, and
    *   its location in the directory tree,
*   diagnose missing or corrupted barcodes (NA rows are explicit),
*   reconcile with external metadata (e.g., from KiwiCloud).

This is vital in breeding and clonal trial contexts where:

*   Audit trails are required,
*   Reprocessing may happen months later,
*   Multiple collaborators may touch the same data.

## 4.5 Step 3 — Decoding barcodes with `resolve_folder_sequence()`

Once your per-directory Excel index files exist, you can decode all images and join results back into those workbooks using `resolve_folder_sequence()`.

```{r resolve-folder-sequence, eval = FALSE}
index_root <- image_root  # or a subfolder containing the index .xlsx files

joined_results <- resolve_folder_sequence(
  path    = index_root,
  pattern = "\\.xlsx$"   # or more specific if needed
)

length(joined_results)
names(joined_results)[1:3]
```

For each workbook, `resolve_folder_sequence()`:

1.  Finds `.xlsx` files under `path` (non-lock files).
2.  Reads the workbook via `readxl::read_xlsx()`.
3.  Checks writeability via `can_write_xlsx()`.
4.  For each row (image), constructs `full_path` and calls `detect_codes_all_zxing()`:
    *   obtains one or more decoded symbols, or
    *   a single NA row if nothing usable is found.
5.  Binds all decoded rows into a long-format tibble.
6.  Applies `resolve_sequence()` to:
    *   parse filename `prefix` and `number`,
    *   statefully propagate `resolved_text`.
7.  Reduces to barcode-specific columns + join key(s).
8.  Performs a **one-to-many join** back onto the original index rows.
9.  Writes the joined table back to the same Excel file via `safe_write_xlsx()`.

The return value is a named list of these joined data frames.

## 4.6 Exploring the joined output

After running `resolve_folder_sequence()`, each Excel workbook now contains:

*   the original per-image index columns, plus
*   barcode-related columns with potentially multiple rows per image.

Inspect the first result:

```{r inspect-joined-output, eval = FALSE}
first_file <- names(joined_results)[1]
first_df   <- joined_results[[1]]

first_file
dplyr::glimpse(first_df)
```

You should see columns such as:

*   `dir`, `subdir`, `file_name`, `ext`, `rel_path`, `full_path` (if present),
*   `index`, `code_format`, `text`, `type`,
*   `prefix`, `number`, `resolved_text`.

Some quick summaries:

```{r summaries, eval = FALSE}
library(dplyr)

# How many images indexed per directory?
first_df %>%
  distinct(dir, file_name) %>%
  count(dir, name = "n_images")

# How many barcode symbols per image?
first_df %>%
  count(file_name, name = "n_symbols") %>%
  arrange(desc(n_symbols)) %>%
  head()

# Most common resolved_text values
first_df %>%
  filter(!is.na(resolved_text)) %>%
  count(resolved_text, sort = TRUE) %>%
  head()
```

## 4.7 Running long scans and writes in the background

On network-mounted drives, `scan_directories()` and `resolve_folder_sequence()` can take **many hours** because:

*   file listing and stat calls are slow,
*   Excel writing is relatively heavy,
*   decoding thousands of images is CPU- and I/O-intensive.

It is often preferable to run these in the background.

### 4.7.1 Using RStudio Jobs

Create a script, e.g. `scripts/run_scan_directories.R`:

```r
# scripts/run_scan_directories.R
library(kiwidecoder)

image_root <- "N:/path/to/your/image/root"  # <-- EDIT

scan_directories(
  root            = image_root,
  extensions      = c("jpg", "jpeg", "png"),
  prefixes        = c("^PMC_[0-9]{5}$", "^DSC_[0-9]{4}$",
                      "^IMG_[0-9]{4}$", "^P[0-9]{7}$"),
  checkpoint_file = "scan_checkpoint.rds",
  resume          = TRUE,
  dry_run         = FALSE,
  log_file        = file.path(image_root, "scan_log.csv"),
  max_path_length = 255
)
```

Then in RStudio:  
**Jobs → Run Script as Job…** → select the script.

Do the same for `resolve_folder_sequence()` if needed.

### 4.7.2 Using `callr::r_bg()` for background execution

```{r background-callr, eval = FALSE}
# install.packages("callr")
library(callr)

image_root <- "N:/path/to/your/image/root"  # <-- EDIT

bg_job <- callr::r_bg(
  func = function(root) {
    library(kiwidecoder)
    scan_directories(
      root            = root,
      extensions      = c("jpg", "jpeg", "png"),
      prefixes        = c("^PMC_[0-9]{5}$", "^DSC_[0-9]{4}$",
                          "^IMG_[0-9]{4}$", "^P[0-9]{7}$"),
      checkpoint_file = "scan_checkpoint.rds",
      resume          = TRUE,
      dry_run         = FALSE,
      log_file        = file.path(root, "scan_log.csv"),
      max_path_length = 255
    )
  },
  args = list(root = image_root)
)

# Monitor / collect result
bg_job$is_alive()
bg_job$get_result()
```

Because the operations are mostly **I/O bound** (network latency), it is perfectly reasonable to let them run in a background process while you continue working interactively.

The `checkpoint_file` ensures you can resume if something interrupts the run (e.g., network hiccups, session timeout).

# 5. Core Decoding and Interpretation Functions

This section describes the key functions that can also be used on their own, outside the full Excel workflow.

## 5.1 `detect_codes_all_zxing()` – decode a single image

`detect_codes_all_zxing(path)` is the **core single-image decoder**.

Behaviour:

*   Normalises the path with `normalizePath()`.
*   Tries to import `PIL.Image` and `zxingcpp`.
*   Opens the image via `PIL.Image.open()`.
*   Calls `zxingcpp.read_barcodes()` with rotation and downscaling enabled.
*   Filters out results where `valid == FALSE` if that field exists.
*   Drops GS1 DataBar formats (not used in this workflow).
*   If nothing usable remains:
    *   returns one NA row with the image path in `file`.
*   Otherwise:
    *   returns one row per symbol with columns: `file`, `index`, `code_format`, `text`, `type`.

Example:

```{r detect-single, eval = FALSE}
img_path <- "N:/Projects/Stage2 Clonal Trials/Images/Male Photos 2024_25/IMG_0001.JPG"

detect_codes_all_zxing(img_path)
```

Typical output:

```text
# A tibble: 3 × 5
  file                               index code_format text    type           
  <chr>                              <int> <chr>       <chr>   <chr>          
1 N:/.../IMG_0001.JPG                    1 QRCode      RS0270  plant-genotype
2 N:/.../IMG_0001.JPG                    2 Code39      MC079   scion         
3 N:/.../IMG_0001.JPG                    3 QRCode      SC-0123 scion         
```

If no usable barcodes are found:

```text
# A tibble: 1 × 5
  file                              index code_format text type
  <chr>                             <int> <chr>       <chr> <chr>
1 N:/.../Hannah Grafting 14 Dec ...    NA NA          NA    NA
```

This **“at least one row per file”** guarantee underpins the traceability design of the whole package.

## 5.2 `decode_hierarchical_zxing()` – hierarchical cropping for difficult images

For some images (e.g. benches with many pots), barcodes may be:

*   small relative to the whole image,
*   off-centre,
*   clustered in particular bands.

`decode_hierarchical_zxing(path)` applies a structured search:

1.  Full image,
2.  Top and bottom halves,
3.  Top/middle/bottom thirds,
4.  3×3 grid tiles (final stage).

At each stage (except the final grid), if **any usable symbols** are found, the function returns those rows and stops.

Each row contains:

*   `file` – original image path,
*   `index`, `code_format`, `text`, `type`,
*   region metadata: `source`, `region_x`, `region_y`, `region_w`, `region_h`.

Example:

```{r decode-hierarchical, eval = FALSE}
img_path <- "N:/Projects/.../Greenhouse/bench_001.JPG"
decode_hierarchical_zxing(img_path)
```

Use this when:

*   ordinary `detect_codes_all_zxing()` returns only NA rows, but you suspect small codes,
*   you want to localise where in the image codes are being detected.

## 5.3 `resolve_sequence()` – stateful interpretation of sequential filenames

`resolve_sequence(df)` takes a **long-format** data frame with at least:

*   `file_name`,
*   `text` (decoded text, possibly `NA`),

and:

*   parses `prefix` + `number` from the filename stem using pattern:

    ```r
    "^([A-Za-z]+)_([0-9]+)$"
    ```

*   collapses rows per file into a per-image `decoded_text` (first non-NA),

*   orders files by `(prefix, number)`,

*   carries forward decoded values along sequential filenames,

*   assigns `"unknown"` when the chain breaks,

*   joins a `resolved_text` column back to each row for that file.

Example:

```{r resolve-sequence-manual, eval = FALSE}
test_df <- tibble::tibble(
  file_name = c("IMG_0001.JPG", "IMG_0001.JPG", "IMG_0002.JPG", "IMG_0004.JPG"),
  text      = c("A1", NA, NA, "B2")
)

resolve_sequence(test_df)
```

Expected `resolved_text`:

```text
[1] "A1" "A1" "A1" "B2"
```

Note:  
In real data, this is applied automatically inside `resolve_folder_sequence()` after decoding all images.

## 5.4 Naming conventions and prefix/number parsing

The sequence logic assumes filenames of the form:

```text
PREFIX_NUMBER.EXT
e.g. IMG_0001.JPG, DSC_0123.JPG
```

If filenames deviate from this pattern:

*   `prefix` and `number` become `NA`,
*   `resolve_sequence()` cannot propagate states across those images,
*   `resolved_text` will often fall back to `"unknown"`.

This is by design:

*   To avoid silently misinterpreting filenames,
*   To make non-conforming images stand out.

In practice, this means:

*   It is helpful to **standardise imaging naming schemes** (e.g., `IMG_0001.JPG`, `IMG_0002.JPG`, …).
*   Files outside these conventions can still be decoded; they simply do not benefit from sequence propagation.

## 5.5 Semantic classification of decoded text (`.classify_type()`)

Internally, `detect_codes_all_zxing()` calls an internal helper (e.g., `.classify_type()`) that:

*   looks at decoded `text`,
*   assigns labels like:
    *   `"plant-genotype"`,
    *   `"scion"`,
    *   `"location"`,
    *   `"scion-genotype"`,
    *   etc.

The exact rules are specific to your naming schemes (e.g., prefixes like `RS`, `MC`, `SC`).

Semantics:

*   `type` is **not required** for the pipeline to work:  
    all joining and sequence logic uses `file_name` and `text`.
*   However, it is extremely useful for:
    *   summarising how many plant vs scion vs location codes,
    *   filtering for only certain types of codes,
    *   QA: checking that the right type of code appears where expected.

If you later change the classification logic, you can re-run:

*   `resolve_folder_sequence()`  
    on existing indexes to refresh `type` labels while preserving the original decodes.

# 6. Diagnostic and Development Helpers

This section covers tools that are **not** part of the main production pipeline but are very useful for:

*   quick tests on small subsets,
*   debugging difficult images,
*   exploring ZXing-CPP behaviour.

## 6.1 `scan_one_folder()` – non-recursive test decoder

`scan_one_folder(folder)` was originally written to:

*   test end-to-end decoding in a **single folder**,
*   avoid waiting for full-tree scans (\~7+ hours on the network),
*   quickly inspect how `detect_codes_all_zxing()` behaves.

It:

*   lists `.jpg`, `.jpeg`, `.png` files in `folder` (non-recursive),
*   calls `detect_codes_all_zxing()` on each,
*   if a given image yields 0 rows (no usable results), injects a single NA row,
*   binds everything into one tibble (one or more rows per file).

Example:

```{r scan-one-folder, eval = FALSE}
test_folder <- "N:/path/to/a/small/test/folder"  # <-- EDIT
res <- scan_one_folder(test_folder)
dplyr::glimpse(res)
```

Typical use cases:

*   sanity-checking the Python/ZXing setup on a small folder,
*   testing the effect of image quality changes,
*   rapid iterations on `.classify_type()` rules.

For **production**, you should use the full index-based pipeline (`scan_directories()` + `resolve_folder_sequence()`), but `scan_one_folder()` remains very handy for development.

## 6.2 Diagnosing difficult images

In practice, not all tags scan cleanly:

*   barcodes can be tiny,
*   tags may be blurred or low contrast,
*   symbols may be detected but undecodable due to damage.

`kiwidecoder` provides two main diagnostic helpers here:

*   `debug_zxing_read()` – inspect raw ZXing-CPP results (including errors),
*   `decode_full_then_center_crop()` – retry decoding on a central crop.

Absolutely — your vignette section *should* be updated now that `debug_zxing_read()` is working cleanly and gives **two distinct kinds of diagnostic outcomes**:

1.  **Case A — ZXing sees *nothing at all* (IMG\_0355.jpg)**
2.  **Case B — ZXing sees multiple symbols, some good, some corrupted (IMG\_6688.jpg)**

You also asked for:

*   clearer, human-readable output (folder, filename separated)
*   improved wording
*   “next steps” guidance
*   updated examples in the vignette

Below is a **fully updated, improved vignette section**, rewritten so you can paste it directly into your Rmarkdown vignette.

It is **clean, accurate, readable**, and reflects the NEW behaviour of your improved `debug_zxing_read()`.

***

### 6.2.1 Inspecting ZXing’s raw output with `debug_zxing_read()`

When `detect_codes_all_zxing()` returns only an NA row, this can mean either:

1. **No barcode‑like region was detected at all**, or  
2. **A barcode was detected but could not be decoded** (e.g. due to damage, blur, occlusion).

These two cases have different practical implications, and  
`debug_zxing_read()` helps you distinguish them.

The function prints **ZXing’s raw per‑symbol diagnostics**, including:

- detected format (`QRCode`, `Code39`, `RMQRCode`, etc.)
- decoded text (if any)
- error messages (e.g., checksum failures)
- validity flag (`TRUE`/`FALSE`)

This makes it possible to tell the difference between  
*“nothing found”* and *“found something but it’s unreadable.”*

---

#### **Case A — No candidate barcode regions found (e.g. IMG_0355.jpg)**

```r
debug_zxing_read("images/IMG_0355.jpg")

Output:

File folder: C:/Users/Kathy.Ruggiero/OneDrive - The Kiwifruit Breeding Centre/Documents/kbc/projects/other/rename-images/images

No candidate symbols or decode errors returned by ZXing for image file: IMG_0355.jpg. 
This usually means ZXing could not find any barcode/QR region at all.

Try the following functions for (possibly) more information:
  - decode_full_then_center_crop(path)
  - decode_hierarchical_zxing(path)

If these do not work, visually inspect the image — the barcode may be 
too small, cropped, low‑contrast, blurred, or otherwise unreadable.
```

**Interpretation:**  
ZXing found *no symbol whatsoever* — not even a damaged one.  
This typically happens when:

*   the barcode is very small in the frame
*   the region is out of focus
*   part of the symbol is cropped
*   lighting/contrast is too poor
*   the symbology is unsupported

In these cases, retaking the photo is usually required.

***

#### **Case B — ZXing detects multiple symbols, with a mix of successes and failures**

```r
debug_zxing_read("images/IMG_6688.jpg")

Output:

    ZXing returned 5 result object(s).

    --- Result 1 ---
     format : BarcodeFormat.QRCode 
     text   : RS022354 
     error  : 
     valid  : TRUE 

    --- Result 2 ---
     format : BarcodeFormat.Code39 
     text   : MC079 
     error  : 
     valid  : TRUE 

    --- Result 3 ---
     format : BarcodeFormat.QRCode 
     text   : SC-002-JWM 
     error  : 
     valid  : TRUE 

    --- Result 4 ---
     format : BarcodeFormat.RMQRCode 
     text   : 
     error  : ChecksumError @ QRDecoder.cpp:361 
     valid  : FALSE 

    --- Result 5 ---
     format : BarcodeFormat.MicroQRCode 
     text   : 82920 
     error  : ChecksumError @ QRDecoder.cpp:361 
     valid  : FALSE

**Interpretation:**

ZXing found five symbol‑like regions:

*   Three decode cleanly (`RS022354`, `MC079`, `SC‑002‑JWM`)
*   Two are mis‑detected tiny QR‑like patterns that fail checksum  
    (`RMQRCode` and `MicroQRCode` with `valid = FALSE`)

Your higher‑level pipeline (`detect_codes_all_zxing()`) already ignores the  
invalid/failed ones and keeps only meaningful, valid results.

This diagnostic display is helpful for:

*   understanding why certain cases return multiple rows
*   verifying that the 1D and QR formats are read correctly
*   confirming that corrupted symbols are being filtered out

***

### When to use `debug_zxing_read()`

Use this function interactively when:

*   you need to understand *why* an image didn’t decode
*   you want to distinguish “no symbol at all” from “symbol damaged”
*   you are developing improvements to your decoding pipeline
*   you want human‑readable insight into ZXing behaviour

Do **not** use it inside bulk workflows — it is a console-level diagnostic tool.
```


### 6.2.2 Retrying on a central crop with `decode_full_then_center_crop()`

In many tray/field photos, the barcode tag occupies only a small central region. `decode_full_then_center_crop()` provides a simple strategy:

1.  Try full image via `detect_codes_all_zxing()`.
2.  If nothing usable, crop the central 50%.
3.  Try decoding the central crop.
4.  Return whichever results are found, with a `source` flag (`"full"` or `"center_crop"`).

Example:

```{r decode-full-center, eval = FALSE}
path <- "N:/Projects/.../Male Photos 2023_24/M6121_5 Dec 23 (1).JPG"

res <- decode_full_then_center_crop(path)
res
```

If the full image decodes:

```text
# A tibble: 2 × 6
  file                             index code_format text    type           source
  <chr>                            <int> <chr>       <chr>   <chr>          <chr> 
1 N:/.../M6121_5 Dec 23 (1).JPG        1 PDF417      RS8401  plant-genotype full  
2 N:/.../M6121_5 Dec 23 (1).JPG        2 QR_CODE     SC1234  scion          full  
```

If only central crop decodes:

```text
# A tibble: 1 × 6
  file                             index code_format text   type    source      
  <chr>                            <int> <chr>       <chr>  <chr>   <chr>       
1 N:/.../M6121_5 Dec 23 (1).JPG        1 DataMatrix  A07    location center_crop
```

If neither decodes, you get a 0-row tibble:

```text
# A tibble: 0 × 6
# ℹ 6 variables: file <chr>, index <int>, code_format <chr>, text <chr>, type <chr>, source <chr>
```

Combined with `debug_zxing_read()`, this gives a powerful toolkit for **explaining why** a particular image fails in the main pipeline.

# 7. Performance and Benchmarking Considerations

`kiwidecoder` is designed for **high-throughput decoding** on **network-mounted drives**, where:

*   I/O is slow,
*   latency is high,
*   lots of small file operations are required.

Key design decisions:

*   **Depth-first scanning** in `scan_directories()`:
    *   completes one subtree at a time,
    *   minimises thrashing between paths,
    *   works well with network caching.

*   **Checkpointing**:
    *   long runs can be resumed (e.g., after 4 hours of work, an interruption only loses a small amount of progress),
    *   allows incremental builds of index files.

*   **Excel index per directory**:
    *   localises write operations,
    *   makes debugging easier (one workbook per folder),
    *   allows partial reprocessing of specific directories.

*   **Atomic Excel writes (`safe_write_xlsx()`)**:
    *   prevents partial files if something goes wrong,
    *   ensures that each workbook is always in a consistent state.

In practice:

*   Full-tree scans plus decoding may take many hours.
*   The main bottlenecks are:
    *   network file listing,
    *   image loading,
    *   Excel I/O,
    *   barcode decoding for large images.

Running these operations in a **background job** is not an optimisation per se, but it greatly improves usability by letting you continue other work while the heavy lifting happens in another process.

# 8. Troubleshooting Python and `reticulate`

This appendix is for users who encounter problems such as:

*   missing Python modules,
*   ZXing not being found,
*   images failing to open,
*   or ambiguous errors.

When the Python environment is correct, decoding should “just work”.

## 8.1 When R cannot find required Python libraries

Errors like:

```text
ModuleNotFoundError: No module named 'PIL'
```

or:

```text
ModuleNotFoundError: No module named 'zxingcpp'
```

mean the Python packages are not installed.

Fix:

```r
reticulate::py_install(
  packages = c("pillow", "zxingcpp"),
  pip = TRUE
)
```

Then restart R and try again.

## 8.2 Multiple Python installations

If `py_config()` shows an unexpected Python path, you may need to choose a specific one:

```r
reticulate::use_python("/path/to/python", required = TRUE)
reticulate::py_install(c("pillow", "zxingcpp"))
```

Then restart R and test again.

## 8.3 Posit Workbench considerations

On Workbench:

*   `py_install()` usually installs packages into a user-private environment.
*   If blocked, ask your admin to install `pillow` and `zxingcpp` into the shared environment.

Once installed where R can see them, `detect_codes_all_zxing()` and `resolve_folder_sequence()` should work normally.

## 8.4 Errors opening images

If you see:

```text
Failed to open image with PIL
```

this may mean:

*   the image file is corrupted,
*   the file is not a supported format,
*   `Pillow` needs reinstalling.

Try:

```r
reticulate::py_install("pillow")
```

Then restart R and test.

## 8.5 When ZXing appears to decode nothing

If all `code_format` values are `NA` even though images clearly contain barcodes:

1.  Check Python config:

    ```r
    reticulate::py_config()
    ```

2.  Test decoding a small folder:

    ```r
    results <- scan_one_folder("path/to/small/test/folder")
    View(results)
    ```

If you only see unsupported formats or NA, it might be that:

*   the barcodes use a format ZXing-CPP does not support (e.g., MicroPDF417),
*   the images are too low-resolution.

In that case, the environment is likely fine; the limitation is on the decoding engine or image quality.

## 8.6 When changes to Python are not recognised

`reticulate` initialises Python once per R session. After any significant changes (new packages, changed environment), restart R:

```r
.rs.restartR()   # in RStudio
```

## 8.7 When to seek further help

If after:

*   installing `pillow` and `zxingcpp`,
*   selecting the correct Python,
*   restarting R,

you still see issues, you may need help from infrastructure support.

Provide:

*   output of `reticulate::py_config()`,
*   error messages,
*   any relevant logs (e.g., long path warnings, decoding errors in the console).

# 9. Diagnostic Script: `diagnose_barcode_python()`

To make troubleshooting easier, you can define a helper function:

```{r diagnose-python, eval = FALSE}
diagnose_barcode_python <- function() {
  cat("=== Barcode Python Environment Diagnostic ===\n\n")
  
  # 1. Check that reticulate is installed
  if (!requireNamespace("reticulate", quietly = TRUE)) {
    cat("reticulate is NOT installed.\n")
    cat("Please install it with:\n\n")
    cat('  install.packages("reticulate")\n\n')
    cat("Then restart R and run this diagnostic again.\n")
    return(invisible(list(
      ok          = FALSE,
      has_retic   = FALSE,
      py_config   = NULL,
      has_pillow  = NA,
      has_zxing   = NA
    )))
  }
  
  cat("reticulate is installed.\n\n")
  
  # 2. Show Python configuration
  cat("Checking Python configuration (reticulate::py_config())...\n\n")
  py_conf <- NULL
  py_conf <- tryCatch(
    reticulate::py_config(),
    error = function(e) {
      cat("Could not obtain Python configuration.\n")
      cat("Error was:\n")
      cat("  ", conditionMessage(e), "\n\n")
      NULL
    }
  )
  
  if (is.null(py_conf)) {
    cat("R could not initialise Python.\n")
    cat("You may need to install Python or configure reticulate with use_python()/use_condaenv().\n")
    return(invisible(list(
      ok          = FALSE,
      has_retic   = TRUE,
      py_config   = NULL,
      has_pillow  = NA,
      has_zxing   = NA
    )))
  }
  
  cat("Python configuration detected:\n")
  cat("  Python path   :", py_conf$python,  "\n")
  cat("  Python version:", py_conf$version, "\n\n")
  
  # 3. Check for Pillow (PIL)
  cat("Checking for Python package 'pillow' (PIL)...\n")
  has_pillow <- reticulate::py_module_available("PIL")
  
  if (!has_pillow) {
    cat("Pillow (PIL) is NOT available in this Python environment.\n")
    cat("You can install it from R with:\n\n")
    cat('  reticulate::py_install("pillow")\n\n')
  } else {
    cat("Pillow (PIL) is available.\n\n")
  }
  
  # 4. Check for zxingcpp
  cat("Checking for Python package 'zxingcpp'...\n")
  has_zxing <- reticulate::py_module_available("zxingcpp")
  
  if (!has_zxing) {
    cat("zxingcpp is NOT available in this Python environment.\n")
    cat("You can install it from R with:\n\n")
    cat('  reticulate::py_install("zxingcpp")\n\n')
  } else {
    cat("zxingcpp is available.\n\n")
  }
  
  # 5. Final summary
  ok <- isTRUE(has_pillow) && isTRUE(has_zxing)
  
  if (ok) {
    cat("✅ Your Python environment appears to be correctly configured for barcode decoding.\n")
    cat("You should now be able to run:\n\n")
    cat('  detect_codes_all_zxing("path/to/image.jpg")\n')
    cat('  scan_one_folder("path/to/folder")\n\n')
  } else {
    cat("⚠ Your Python environment is NOT fully configured yet.\n")
    cat("Please install any missing packages as indicated above, restart R, and run this diagnostic again.\n\n")
  }
  
  invisible(list(
    ok          = ok,
    has_retic   = TRUE,
    py_config   = py_conf,
    has_pillow  = has_pillow,
    has_zxing   = has_zxing
  ))
}
```

Run:

```r
diagnose_barcode_python()
```

to see:

*   which Python is being used,
*   whether `pillow` and `zxingcpp` are available,
*   a simple pass/fail summary.


